Illusion Game & PCS

### 2025-04-28 analysis/1_cleaning.qmd
changes to anlaysis scripts

### PCS

pcs <-  select(data_ppt, starts_with("pcs_"))

pcs$pcs_tastescore <- rowMeans(pcs[grepl("taste", names(pcs))])

pcs$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)

pcs$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)

data_ppt$pcs <- rowMeans(pcs[!grepl("pcs_amnesia_w", names(pcs)) & grepl("hands|music|mosquito|tastescore|arm|negative|postsession|amnesia", names(pcs))], na.rm = TRUE)

data_ppt$pcs_SD <- sd(data_ppt$pcs, na.rm = TRUE)
```

@@

@@

>>>>

@@

**Note**. If the RT is collapsed to the left (i.e., has > 1/3 of ultra-fast responses - typically < 200 ms) in the first set, the entire participant will be discarded (suggesting that they did not properly do the task), but if only the second set is bad, then only the second set will be discarded (as the illusion sensitivity can still be estimated, albeit with less precision).

@@

**Note**. The removal of individual trials will also be performed [RT < 200 ms or > 3 SD; following Thériault et al. (2024)]. To mitigate the risk of confounding effects driven by extreme speed or accuracy strategies, participants whose RTs are significantly slower than the group average (RT > 4 SD above the mean, based on Makowski et al., 2023) will be excluded from the analysis.

### 2025-04-28 analysis/2_analysis.qmd
changes to anlaysis scripts

library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(brms)

@@

df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/data_participants.csv") |>
  mutate(
    Block = as.factor(Block),
    Illusion_Side = as.factor(Illusion_Side),
    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent")
  ) 
```

@@

@@

>>>>

library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(psych)

@@

# data_ppt <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_participants.csv")

data_ppt <- read.csv("C:/Users/asf25/Desktop/studies/IllusionGamePhenomenologicalControl/data/rawdata_participants.csv")

# df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/data_participants.csv")  |>
  # group_by(Illusion_Type) |> 
  # mutate(Illusion_Side = sign(Illusion_Difference), 
  #   Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
  #   Task_Difficulty = abs(Illusion_Difference),
  #        Condition_Illusion = datawizard::categorize(
  #          Illusion_Strength, split="quantile", n_groups=4,
  #          labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
  #        Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
  # ungroup() |>
  # mutate(
  #   Block = as.factor(Block),
  #   Illusion_Side = as.factor(Illusion_Side),
  #   Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent")
  # ) 
  #

#taken from illusion game international
data_ig <- read.csv("../data/data_IllusionGame.csv") |>
  group_by(Illusion_Type) |> 
  mutate(Illusion_Side = sign(Illusion_Difference), 
    Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
    Task_Difficulty = abs(Illusion_Difference),
      Condition_Illusion = case_when(
        Illusion_Effect == "Congruent" ~ "Congruent",
        Illusion_Effect == "Incongruent" & Illusion_Strength <= median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Mild",
        Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"),
    Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
  ungroup() |>
  mutate(
    Block = as.factor(Block),
    Illusion_Side = as.factor(Illusion_Side),
    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
    Response_Correct = as.factor(Response_Correct)
  ) 
```

@@

*Note:* Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score. 
The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
The subjective score for item 10 is the geometric mean (the square root of the product) of he responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion.

```{r}
data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
data_ppt$pcs <- rowMeans(data_ppt[, c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")], na.rm = TRUE)
data_ppt$pcs_SD <- sd(data_ppt$pcs, na.rm = TRUE)

### Reliability
```{r}

data_ppt |>
  select("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia") |>
  psych::alpha()

### Error rate

**Note.** Error rate computed by dividing  the number of errors by the total number of trials.

Error_rate <- data_ig |>
  group_by(Participant, Illusion_Type, Condition_Illusion) |>
  summarise(
    incorrect_responses = sum(Response_Correct == "false", na.rm = TRUE),
    total_responses = n(),
    error_rate = (incorrect_responses / total_responses) * 100
  )

Error_rate

#### Difference in Error rate between each condition and illusion type

::: {.panel-tabset}

##### Ebbinghaus

ER_ebbinghaus <- Error_rate |> filter(Illusion_Type == "Ebbinghaus")

ER_CongruentMild_ebbinghaus_BF <- BayesFactor::ttestBF(
  x = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Congruent"],
  y = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Mild"],
  paired = TRUE
)

ER_StrongMild_ebbinghaus_BF <- BayesFactor::ttestBF(
  x = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Mild"],
  y = ER_ebbinghaus$error_rate[ER_ebbinghaus$Condition_Illusion == "Strong"],
  paired = TRUE
)

##### MullerLyer

ER_mullerlyer <- Error_rate |> filter(Illusion_Type == "MullerLyer")

ER_CongruentMild_mullerlye_BF <- BayesFactor::ttestBF(
  x = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Congruent"],
  y = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Mild"],
  paired = TRUE
)

ER_StrongMild_mullerlye_BF <- BayesFactor::ttestBF(
  x = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Mild"],
  y = ER_mullerlyer$error_rate[ER_mullerlyer$Condition_Illusion == "Strong"],
  paired = TRUE
)

##### Vertical-Horizontal

ER_verticalhorizontal  <- Error_rate |> filter(Illusion_Type == "VerticalHorizontal")

ER_CongruentMild_vh_BF <- BayesFactor::ttestBF(
  x = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Congruent"],
  y = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Mild"],
  paired = TRUE
)

ER_StrongMild_vh_BF <- BayesFactor::ttestBF(
  x = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Mild"],
  y = ER_verticalhorizontal$error_rate[ER_verticalhorizontal$Condition_Illusion == "Strong"],
  paired = TRUE
)

:::

### IES

**Note.** The IES is calculated by dividing the average correct RT by the proportion of correct responses, for each illusion and for each illusion strenght group: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent

IES <- data_ig |>
  group_by(Participant, Illusion_Type, Condition_Illusion) |>
  summarise(
    mean_RT = mean(RT[Response_Correct == "true"], na.rm = TRUE),  # Only calculate mean RT for correct responses
    correct_proportion = mean(Response_Correct == "true", na.rm = TRUE),  # Proportion of correct responses
    IES = mean_RT / correct_proportion  # Compute IES
  )

IES

#### Difference in IES between each condition and illusion type

::: {.panel-tabset}

##### Ebbinghaus

IES_ebbinghaus <- IES |> filter(Illusion_Type == "Ebbinghaus")

BF_Congruent_vs_Mild_ebbinghaus <- BayesFactor::ttestBF(
  x = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Congruent"],
  y = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Mild"],
  paired = TRUE
)

BF_Mild_vs_Strong_ebbinghaus <- BayesFactor::ttestBF(
  x = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Mild"],
  y = IES_ebbinghaus$IES[IES_ebbinghaus$Condition_Illusion == "Strong"],
  paired = TRUE
)

##### MullerLyer

IES_mullerlyer <- IES |> filter(Illusion_Type == "MullerLyer")

BF_Congruent_vs_Mild_mullerlyer <- BayesFactor::ttestBF(
  x = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Congruent"],
  y = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Mild"],
  paired = TRUE
)

BF_Mild_vs_Strong_mullerlyer <- BayesFactor::ttestBF(
  x = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Mild"],
  y = IES_mullerlyer$IES[IES_mullerlyer$Condition_Illusion == "Strong"],
  paired = TRUE
)

##### Vertical-Horizontal

IES_verticalhorizontal  <- IES |> filter(Illusion_Type == "VerticalHorizontal")

BF_Congruent_vs_Mild_verticalhorizontal <- BayesFactor::ttestBF(
  x = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Congruent"],
  y = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Mild"],
  paired = TRUE
)

BF_Mild_vs_Strong_verticalhorizontal <- BayesFactor::ttestBF(
  x = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Mild"],
  y = IES_verticalhorizontal$IES[IES_verticalhorizontal$Condition_Illusion == "Strong"],
  paired = TRUE
)

:::

### Correlations

**Note**. If correlations high between mild and strong conditions then the mild and strong illusion strength groups will be collapsed and the outcomes will be recomputed across all trials, otherwise they will be treated as separate in subsequent analyses.

#### Error Rate

ER_Incongruent_eb <- Error_rate |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "Ebbinghaus")
BayesFactor::correlationBF(
  y = ER_Incongruent_eb$error_rate[ER_Incongruent_eb$Condition_Illusion == "Mild"], 
  x = ER_Incongruent_eb$error_rate[ER_Incongruent_eb$Condition_Illusion == "Strong"])

ER_Incongruent_ml <- Error_rate |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "MullerLyer")
BayesFactor::correlationBF(
  y = ER_Incongruent_ml$error_rate[ER_Incongruent_ml$Condition_Illusion == "Mild"], 
  x = ER_Incongruent_ml$error_rate[ER_Incongruent_ml$Condition_Illusion == "Strong"])

ER_Incongruent_vh <- Error_rate |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "VerticalHorizontal")
BayesFactor::correlationBF(
  y = ER_Incongruent_vh$error_rate[ER_Incongruent_vh$Condition_Illusion == "Mild"], 
  x = ER_Incongruent_vh$error_rate[ER_Incongruent_vh$Condition_Illusion == "Strong"])

#### IES

IES_Incongruent_eb <- IES |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "Ebbinghaus")
BayesFactor::correlationBF(
  y = IES_Incongruent_eb$IES[IES_Incongruent_eb$Condition_Illusion == "Mild"], 
  x = IES_Incongruent_eb$IES[IES_Incongruent_eb$Condition_Illusion == "Strong"])

IES_Incongruent_ml <- IES |> filter(Condition_Illusion != "Congruent" & Illusion_Type == "MullerLyer")
BayesFactor::correlationBF(
  y = IES_Incongruent_ml$IES[IES_Incongruent_ml$Condition_Illusion == "Mild"], 
  x = IES_Incongruent_ml$IES[IES_Incongruent_ml$Condition_Illusion == "Strong"])

IES_Incongruent_vh <- IES|> filter(Condition_Illusion != "Congruent" & Illusion_Type == "VerticalHorizontal")
BayesFactor::correlationBF(
  y = IES_Incongruent_vh$IES[IES_Incongruent_vh$Condition_Illusion == "Mild"], 
  x = IES_Incongruent_vh$IES[IES_Incongruent_vh$Condition_Illusion == "Strong"])

*Note*. Reliability analyses are conducted on all resulting indices.

#### Split-half

@@

#### Inter-illusion reliability
